\section*{A Tool to Simplify Network Administration in the Modern Home}

\subsection*{Problem}

As computing becomes more and more ubiquitous, home networks are becoming
increasingly complex. In the future we may see households containing a very
large number of devices, performing functions we can't yet even imagine. 

As home networks grow in both number of hosts and bandwidth requirements, the
task of administering such networks will inevitably become more complicated. 

Unfortunately, most homeowners don't have degrees in computer science, so will
find it hard to set up effective networks in their homes. 

\subsection*{Goals}

The project's main aim is to simplify the administration of complex home
networks. This goal shall be realised by providing users with a system that
either reduces the complexity of managing a complex, multi-subnet, home
network, or providing tools for monitoring the activity of the network. 

\subsection*{Scope}


The project will involve either:
\begin{itemize}
\item Implementing a zeroconf OSPF routing system based on an open source
  routing implementation such as Quagga or XORP.

\item Create a traffic monitoring system that can report on bandwidth usage,
  network flows and perform some form of malware detection. With the
  possibility of implementing a front-end to allow easy QoS configuration. 

\end{itemize}

Both of these projects will involve the use of lightweight Linux boxes to
perform their tasks. The project that will be pursued will be decided upon by
week four. 
